---
title: 'Use BiDAG-package with user-defined score'
author: "Vera Kvisgaard"
date: "2024-04-25"
output: 
  html_document:
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	results = "hold"
)
```

```{r}
# prep
here::i_am("./rmd/backdoor_estimates.Rmd")
files <- list.files(here::here("R"), ".R", full.names = T)
sapply(files, source)
```



Load network and sample data
```{r}

dag <- rbind(L  = c(0, 0, 1, 0, 1),
             Z1 = c(0, 0, 0, 1, 0),
             Z2 = c(0, 0, 0, 1, 0),
             X  = c(0, 0, 0, 0, 1),
             Y  = rep(0, 5))
colnames(dag) <- rownames(dag)
nlev <- rep(2, ncol(dag))
Rgraphviz::plot(as(dag, "graphNEL"))

labels <- vector("list", ncol(dag))
labels[[4]] <- list(NULL, 0)
labels[[5]] <- list(1, NULL) 

bn <- rand_bn(dag, labels, nlev)
bn
```

```{r}


x  <- 4
yvars <- 5
P <- labels_to_partition(labels[[4]], nlev[z])
pa <- 2:3
pa_min <- 3
q <- prod(nlev[pa])
    
dims <- list(iter = 1:10, 
             N = 10**c(1:4))
rmse <- array(list(), dim = lengths(dims), dimnames = dims)
kl   <- function(p, q) sum(p*(log(p)-log(q)))
for (i in seq_along(dims$iter)) {
  set.seed(i)
  bn <- rand_bn(dag, labels, nlev)
  pdo  <- bida:::interv_probs_x_bn(bn, x, yvars, bida:::descendants(bn))[[1]]
  pyxz <- bida:::cpquery_from_cpt_arrays(bida:::cpt_arrays_from_bn(bn), yvars, c(x, pa))
  diff <- pdo[2, 2]-pdo[1, 2]
  for (j in seq_along(dims$N)) {
  
    N <- dims$N[j]
    set.seed(i+N)
    data <- bida:::sample_data_from_bn(bn, N)

    pairs <- list(part = bida_pairs_local_csi(data, rbind(pa), 1, x, yvars, P, ess = 1, kappa = .5)[[1]],
                  full = bida_pairs_local_csi(data, rbind(pa), 1, x, yvars, as.list(seq_len(q)-1), ess = 1, kappa = .5)[[1]],
                  min  = bida_pairs_local_csi(data, rbind(pa_min), 1, x, yvars, NULL, ess = 1, kappa = .5)[[1]])
    
    means_obs <- lapply(lapply(pairs, "[[", "params"),
                        function(x) sweep(x[[1]]+1/length(x[[1]]), 2:3, colSums(x[[1]])+nlev[yvars]/length(x[[1]]),  "/")) 
    
    means <- lapply(pairs, bida:::posterior_mean)
    rmse[[i, j]]  <- sapply(means, function(x) sqrt(mean((x-pdo)**2)))
    #rmse[[i, j]] <- sapply(means, function(x) kl(pdo, x))  
  }
}


library(ggplot2)
library(dplyr)
df <- cbind(expand.grid(dims), do.call("rbind", rmse))
df %>% 
  tidyr::pivot_longer(names(pairs)) %>% 
  mutate(N = factor(N, sort(unique(N)))) %>% 
  ggplot(aes(N, value, color = name)) +
  #facet_wrap(N~., scales = "free") +
  geom_boxplot()
```


Run partition MCMC to obtain a sample of parent sets for each node. 

Define a `scoreparameters`-object for `BiDAG`'s MCMC functions.
Note that the level of each variable is added to the list of parameters. 
I think this argument is used in the PC-algorithm-step, but it is not added when `type = "usr"`.
```{r}
library(BiDAG)
scorepar <- scoreparameters("usr", 
                         data = data,
                         usrpar = list(pctesttype = "bdecat"))
scorepar$Cvec <- nlev 
```

Define the user-specified scoring function that is to be called from `BiDAG`-functions.
Also, assign this function to the name-space of `BiDAG`-package, as otherwise an internal function is called. 
```{r,}
#' See `scoretype` parameter in [BiDAG::scoreparameters].
usrDAGcorescore <- function(j, parentnodes, n, param) {
  local_CSI_score(param$data, nlev, j, parentnodes, ess = 1, kappa = .5)
}

assignInNamespace("usrDAGcorescore", usrDAGcorescore, ns = "BiDAG")
```
```

```{r}

ps <- bida:::parent_support_from_dags()
```


