---
title: 'Run approx BIDA with labeled DAGs'
author: "Vera Kvisgaard"
output: 
  html_document:
  toc: true
---

```{r}

library(ggplot2)
library(dplyr)
library(pcalg)
library(BiDAG)


here::i_am("./rmd/run_BIDA_with_LDAGs.Rmd")
# source functions from file
files <- list.files(here::here("R"), ".R", full.names = T)
sapply(files, source)

verbose <- FALSE
runMCMC <- FALSE
```

Define underlying labeled DAG
```{r}
n <- 10
nlev <- rep(2, n)
dag <- matrix(0, n, n)
colnames(dag) <- rownames(dag) <- paste0("X", seq_len(n))


dag[1, 2:5] <- 1
dag[2, c(3, 7)] <- 1
dag[3, c(4, 7)] <- 1
dag[4, c(5, 7, 8)] <- 1
dag[5, c(9)] <- 1
dag[6, c(7, 10)] <- 1
dag[7, c(8, 10)] <- 1
dag[8, c(5, 9, 10)] <- 1
dag[9, c(10)] <- 1

labels <- matrix(list(), n, n)
labels[[2, 3]] <- rbind(0)
labels[[1, 4]] <- rbind(1)
labels[[4, 5]] <- rbind(c(0, NA))
labels[[8, 5]] <- rbind(c(0, NA))
labels[[2, 7]] <- rbind(c(1, 1, 0))
labels[[3, 7]] <- rbind(c(0, 1, 1), c(1, NA, 1))
labels[[4, 7]] <- rbind(c(1, 1, NA))
labels[[6, 7]] <- rbind(c(1, 1, NA))
labels[[5, 9]] <- rbind(1)
labels[[7, 10]] <- rbind(c(1, NA, NA))
labels[[8, 10]] <- rbind(c(1, NA, NA))
labels[[9, 10]] <- rbind(c(1, NA, NA))


g <- as(dag, "graphNEL")

indx <- sapply(labels[dag == 1], is.null)
prettify_lab <- function(x) {
  tmp <- apply(x, 1, function(y) sprintf("(%s)",  paste(y, collapse = ",")))
  sprintf("{%s}", paste(gsub("NA", "*", tmp), collapse = ","))
}

edgeAttr <- sapply(labels[dag == 1][!indx], prettify_lab)
names(edgeAttr) <- Rgraphviz::edgeNames(g)[!indx]
Rgraphviz::plot(g, edgeAttrs = list(label = edgeAttr))

# store labels as nested list
labels <- lapply(seq_len(n), function(i) labels[which(dag[, i] == 1), i])

# draw distribution over labeled DAG
set.seed(007)
bn <- rand_bn(dag, labels, nlev, alpha = 1)
pdos <- bida:::interv_probs_from_bn(bn, "exact")
tau  <- vapply(pdos, bida:::avg_jsd_array, numeric(1))

dmat <- bida:::descendants(dag)
dindx <- diag(n) == 1
strong_effects  <- tau[!dindx] > quantile(tau[!dindx][dmat[!dindx]==1], .8)


```


Specify a user-defined score function for labeled DAGs, to be used with `BiDAG::scoreparameters` 
Also, assign this function to the name-space of `BiDAG`-package, as otherwise an internal function is called. 
```{r,}
ess <- 1
kappa <- .25

#' See `scoretype` parameter in [BiDAG::scoreparameters].
usrDAGcorescore <- function(j, parentnodes, n, param) {
  local_CSI_score(param$data, nlev, j, parentnodes, ess = ess, kappa = kappa)
}
assignInNamespace("usrDAGcorescore", usrDAGcorescore, ns = "BiDAG")

```

Run partition MCMC and write to file.
```{r}
dims <- list(r = 1:30,
             N = c(300, 1000, 3000),
             model = c("DAG", "LDAG", "KNOWN"))
if (runMCMC) {
  for (r in dims$r) {
  for (NN in seq_along(dims$N)) {
    N <- dims$N[NN]
    set.seed(r+N)
    data <- bida:::sample_data_from_bn(bn, N)
    
    for (m in dims$model) {
      if (verbose) cat(sprintf("\n iteration: %s, N: %s, model: %s\n", r, N, m))
    
      
      if (m == "DAG") {
        scorepar <- scoreparameters("bdecat", data = as.data.frame(data), bdecatpar = list(chi = 1, edgepf = 1))
      } else {
        scorepar <- scoreparameters("usr", data = data, usrpar = list(pctesttype = "bdecat"))
        scorepar$Cvec <- nlev
      }
      tic <- Sys.time()
      iterfit  <- learnBN(scorepar, "orderIter", verbose = TRUE, scoreout = TRUE, hardlimit = 5)
      print(Sys.time()-tic)
      smpl     <- sampleBN(scorepar, "partition", scoretable = getSpace(iterfit))
      print(Sys.time()-tic)
      
      filename <- sprintf("%s_N%s_r%02.0f.rds", m, N, r)
      saveRDS(list(bn, iterfit, smpl), here::here("./MCMCchains/", filename))
      
    }
  }
}
}

```

```{r}

if (runMCMC) {
  nlev <- vapply(bn, function(x) dim(x$prob)[1], integer(1))
dims <- list(r = 11:30,
             N = c(300, 1000, 3000),
             model = c("DAG", "LDAG", "KNOWN"))
  for (r in dims$r) {
  for (NN in seq_along(dims$N)) {
    N <- dims$N[NN]
    set.seed(r+N)
    data <- bida:::sample_data_from_bn(bn, N)
    
    for (m in dims$model) {
      if (verbose) cat(sprintf("\n iteration: %s, N: %s, model: %s\n", r, N, m))
    
      
      if (m == "DAG") {
        scorepar <- scoreparameters("bdecat", data = as.data.frame(data), bdecatpar = list(chi = 1, edgepf = 1))
      } else {
        scorepar <- scoreparameters("usr", data = data, usrpar = list(pctesttype = "bdecat"))
        scorepar$Cvec <- nlev
      }
      tic <- Sys.time()
      iterfit  <- learnBN(scorepar, "orderIter", verbose = TRUE, scoreout = TRUE, hardlimit = 5)
      print(Sys.time()-tic)
      smpl     <- sampleBN(scorepar, "partition", scoretable = getSpace(iterfit))
      print(Sys.time()-tic)
      
      filename <- sprintf("asia_%s_N%s_r%02.0f.rds", m, N, r)
      saveRDS(list(bn, iterfit, smpl), here::here("./MCMCchains/", filename))
      
    }
  }
}
}
```

Load sampled DAGs from file and compute intervention distributions
```{r}
# discard first 200 DAGs
burnin <- seq_len(200)          

# init structures for storing sim results
mse <- taus <- edgep <- ancp <- array(list(), lengths(dims), dims)

for (r in dims$r) {
  for (NN in seq_along(dims$N)) {
    N <- dims$N[NN]
    set.seed(r+N)
    data <- bida:::sample_data_from_bn(bn, N)
    

    for (m in dims$model) {
      if (verbose) cat(sprintf("\n iteration: %s, N: %s, model: %s\n", r, N, m))
      
      # compute support over parent sets
      if (m == "KNOWN") {
        ps   <- bida:::parent_support_from_dags(list(dag), 1)
      } else {
        # load MCMC-chain object from file
        filename <- sprintf("%s_N%s_r%02.0f.rds", m, N, r)
        smpl <- readRDS(here::here("./MCMCchains/", filename))[[3]]
        
        # list unique DAGs
        dags <- lapply(smpl$traceadd$incidence[-burnin], as.matrix)
        u    <- unique(dags)
        support  <- bida:::rowsum_fast(rep(1/length(dags), length(dags)), dags, u)
        dags <- u 
        dmats <- lapply(dags, bida:::descendants)
        
        # compute edge prob and ARP
        edgep[[r, NN, m]] <- Reduce("+", Map("*", dags, support))
        ancp[[r, NN, m]]  <- Reduce("+", Map("*", dmats, support))
        
        ps   <- bida:::parent_support_from_dags(dags, support)
      }
     

      
      # compute MSE of intervention distrib params
      bida <- bida:::bida(ps, data, "categorical", list(nlev = nlev, ess = ess))
      means <- bida:::posterior_mean(bida)
      tmp <- vapply(seq_along(means)[!dindx], 
                    function(x) mean((means[[x]]-pdos[[x]])**2), numeric(1))
      mse[[r, NN, m]] <- tmp
      taus[[r, NN, m]] <- vapply(means[!dindx], bida:::avg_jsd_array, numeric(1))
    }
  }
}
```

### MSE of intervention distribution params
```{r}
df <- cbind(expand.grid(dims)[!sapply(mse, is.null), ], do.call("rbind", mse))
df <- cbind(expand.grid(dims)[!sapply(mse, is.null), ], sqrt(vapply(mse, mean, numeric(1))))

df %>% 
  tidyr::pivot_longer(-c(1:3)) %>% 
  mutate(N = factor(N, sort(unique(N)))) %>% 
  ggplot(aes(N, value, color = model)) +
  #facet_grid(, scales = "free") +
  theme_minimal() +
  theme(legend.position = "top",
        legend.title=element_blank()) +
  geom_boxplot() +
  ylab("Root-squared mean errors (RMSE)") +
  xlab("Sample size") 
```

### Precision recall
```{r}
compute_prec_recall <- function(x, y) {
  indx <- order(x, decreasing = TRUE)
  tp  <- cumsum(y[indx])
  cbind(indx = indx, TPR = tp/sum(y), PPV = tp/seq_along(y))
}

# edge probs
par <- expand.grid(dimnames(edgep[,,-3]))
tmp <- lapply(edgep[,,-3], function(x) compute_prec_recall(unlist(x), dag))
indx <- rep(seq_len(nrow(par)), each = n**2)
df  <- cbind(par[indx, ], do.call("rbind", tmp))

df %>% 
  ggplot(aes(TPR, PPV, color = r)) +
  facet_grid(N~model) +
  geom_hline(aes(yintercept = mean(dag)), size = .2) +
  geom_line() +
  ylab("Precision") +
  xlab("True positive rate") +
  ggtitle("Edge probability") 

# ancestor relation probs
par <- expand.grid(dimnames(ancp[,,-3]))
tmp <- lapply(ancp[,,-3], function(x) compute_prec_recall(unlist(x), dmat))
indx <- rep(seq_len(nrow(par)), each = n**2)
df  <- cbind(par[indx, ], do.call("rbind", tmp))

df %>% 
  ggplot(aes(TPR, PPV, color = r)) +
  facet_grid(N~model) +
  geom_hline(aes(yintercept = mean(dmat)), size = .2) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("Precision") +
  xlab("True positive rate") 

# Strong effects
par <- expand.grid(dimnames(taus))
tmp <- lapply(taus, function(x) compute_prec_recall(unlist(x), strong_effects))
indx <- rep(seq_len(nrow(par)), each = n*(n-1))
df  <- cbind(par[indx, ], do.call("rbind", tmp))

df %>% 
  ggplot(aes(TPR, PPV, color = r)) +
  facet_grid(N~model) +
  geom_hline(aes(yintercept = mean(strong_effects)), size = .2) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("Precision") +
  xlab("True positive rate") 
#  ggtitle("Top 10th percentile strongest effect")

```

```{r}

# avg ppv 
avgppv <- function(x, which_true) {
  pp  <- (length(x)-rank(x) +1)[which_true]
  ppv <- seq_along(pp)/sort(pp)
  mean(ppv)
}

tmp <- lapply(taus, function(x) avgppv(x, which(strong_effects == 1)))
df  <- cbind(expand.grid(dimnames(taus)), value = do.call("rbind", tmp)) 


df %>% 
  group_by(N, model) %>% 
  summarize(mean = mean(value),
            sd   = sd(value)) %>% 
  tidyr::pivot_longer(c("mean", "sd")) %>% 
  mutate(value = as.character(round(value, 2)), 
         value = ifelse(name == "sd", paste0("(", value, ")"), value)) %>% 
  tidyr::pivot_wider(names_from = "model") -> tab

tab %>% 
  select(-name) %>% 
  knitr::kable(digits = c(2))


# positive effects
tmp <- lapply(taus, function(x) avgppv(x, which(dmat[!dindx] == 1)))
df  <- cbind(expand.grid(dimnames(taus)), value = do.call("rbind", tmp)) 


df %>% 
  group_by(N, model) %>% 
  summarize(mean = mean(value),
            sd   = sd(value)) %>% 
  tidyr::pivot_longer(c("mean", "sd")) %>% 
  mutate(value = as.character(round(value, 2)), 
         value = ifelse(name == "sd", paste0("(", value, ")"), value)) %>% 
  tidyr::pivot_wider(names_from = "model") -> tab

tab %>% 
  select(-name) %>% 
  knitr::kable(digits = c(2))

```

