---
title: 'Run approx BIDA with labeled DAGs'
author: "Vera Kvisgaard"
output: 
  html_document:
    toc: true
---

```{r message=FALSE, warning=FALSE}

library(ggplot2)
library(dplyr)
library(pcalg)
library(BiDAG)
library(Rgraphviz)
library(pcalg)

here::i_am("./rmd/run_BIDA_with_LDAGs.Rmd")
# source functions from file
files <- list.files(here::here("R"), ".R", full.names = T)
invisible(sapply(files, source))

verbose <- FALSE
runMCMC <- FALSE
```

### Draw a random Bayesian network
Define labeled DAG
```{r}

dag <- rbind(L  = c(0, 1, 1, 0, 1),
             Z1 = c(0, 0, 0, 1, 0),
             Z2 = c(0, 0, 0, 1, 0),
             X  = c(0, 0, 0, 0, 1),
             Y  = rep(0, 5))
colnames(dag) <- rownames(dag)
n <- ncol(dag)
nlev <- rep(2, n)

labels <- matrix(list(), n, n)
labels[[2, 4]] <- rbind(0)  # Z1 -> X
labels[[3, 4]] <- rbind(0)  # Z2 -> X
labels[[4, 5]] <- rbind(0)  # X -> Y

# plot
g <- as(dag, "graphNEL")
indx <- sapply(labels[dag == 1], is.null)
prettify_lab <- function(x) {
  tmp <- apply(x, 1, function(y) sprintf("(%s)",  paste(y, collapse = ",")))
  sprintf("{%s}", paste(gsub("NA", "*", tmp), collapse = ","))
}

lab <- sapply(labels[dag == 1][!indx], prettify_lab)
fonts <- rep(10, length(lab))
names(lab) <- names(fonts) <- Rgraphviz::edgeNames(g)[!indx]
Rgraphviz::plot(g, edgeAttrs = list(label = lab, fontsize = fonts))


# store labels as nested list
labels <- lapply(seq_len(n), function(i) labels[which(dag[, i] == 1), i])

# draw distribution over labeled DAG
set.seed(007)
bn <- rand_bn(dag, labels, nlev, alpha = 1)
pdos <- bida:::interv_probs_from_bn(bn, "exact")
tau  <- vapply(pdos, bida:::avg_jsd_array, numeric(1))

dmat <- bida:::descendants(dag)
dindx <- diag(n) == 1
strong_effects  <- tau[!dindx] > quantile(tau[!dindx][dmat[!dindx]==1], .8)
```
### Score function
Define a score-function that can be used with `BiDAG`. 
Need to assign this function to the namespace of the package, otherwise an internal function is called.
```{r}

# wrapper function
usrDAGcorescore <- function(j, parentnodes, n, scorepar) {
  score_from_lookup(scorepar$data, scorepar$Cvec, j, parentnodes, scorepar$tab)
}

# assign to namespace
assignInNamespace("usrDAGcorescore", usrDAGcorescore, ns = "BiDAG")

```


Specify a `BiDAG::scoreparameters` for LDAGs and standard (discrete) DAGs.
```{r}
get_scorepar <- function(m, data, nlev, tab = NULL) {
 if (m == "DAG") {
      scorepar <- scoreparameters("bdecat", data = as.data.frame(data), bdecatpar = list(chi = 1, edgepf = 1))
    } else {
      if (is.null(tab)) {
        tab <- init_lookup_scoretable(ncol(data), list(ess = 1, kappa = .5))
      }
      scorepar <- scoreparameters("usr", data = data, usrpar = list(pctesttype = "bdecat"))
      scorepar$Cvec <- nlev
      scorepar$tab <- tab
    }
  return(scorepar)
}
```

### Run partition MCMC 
Run partition MCMC to sample DAGs and write results to file.
```{r}
dims <- list(r = 1:10,
             N = c(100, 300, 1000),
             model = c("DAG", "LDAG", "KNOWN"))
if (runMCMC) {
  for (r in dims$r) {
    for (NN in seq_along(dims$N)) {
      N <- dims$N[NN]
      set.seed(r+N)
      data <- bida:::sample_data_from_bn(bn, N)
      
      for (m in dims$model) {
        if (verbose) cat(sprintf("\n iteration: %s, N: %s, model: %s\n", r, N, m))
        if (m == "KNOWN") next
        filename <- sprintf("LDAG5_nodes_%s_N%s_r%02.0f.rds", m, N, r)
        if (file.exists(here::here("./MCMCchains/", filename))) next
        
        scorepar <- get_scorepar(m, data, nlev)
        tic <- Sys.time()
        iterfit  <- learnBN(scorepar, "orderIter", verbose = TRUE, scoreout = TRUE, hardlimit = 5)
        print(Sys.time()-tic)
        smpl     <- sampleBN(scorepar, "partition", scoretable = getSpace(iterfit))
        print(Sys.time()-tic)
        
        saveRDS(list(bn, iterfit, smpl, scorepar), here::here("./MCMCchains/", filename))
        
      }
    }
  }
}

```

### Compute causal params
Load sampled DAGs from file and compute intervention distributions
```{r}
# discard first 200 DAGs
burnin <- seq_len(200)          

# init structures for storing sim results
mse <- taus <- edgep <- ancp <- array(list(), lengths(dims), dims)

for (r in dims$r) {
  for (NN in seq_along(dims$N)) {
    N <- dims$N[NN]
    set.seed(r+N)
    data <- bida:::sample_data_from_bn(bn, N)
    

    for (m in dims$model) {
      if (verbose) cat(sprintf("\n iteration: %s, N: %s, model: %s\n", r, N, m))
      
      # compute support over parent sets
      if (m == "KNOWN") {
        ps   <- bida:::parent_support_from_dags(list(dag), 1)
      } else {
        # load MCMC-chain object from file
        filename <- sprintf("LDAG5_nodes_%s_N%s_r%02.0f.rds", m, N, r)
        if (!file.exists(here::here("./MCMCchains/", filename))) next
        smpl <- readRDS(here::here("./MCMCchains/", filename))[[3]]
        
        # list unique DAGs
        dags <- lapply(smpl$traceadd$incidence[-burnin], as.matrix)
        u    <- unique(dags)
        support  <- bida:::rowsum_fast(rep(1/length(dags), length(dags)), dags, u)
        dags <- u 
        dmats <- lapply(dags, bida:::descendants)
        
        # compute edge prob and ARP
        edgep[[r, NN, m]] <- Reduce("+", Map("*", dags, support))
        ancp[[r, NN, m]]  <- Reduce("+", Map("*", dmats, support))
        
        ps   <- bida:::parent_support_from_dags(dags, support)
      }
     

      
      # compute MSE of intervention distrib params
      bida <- bida:::bida(ps, data, "categorical", list(nlev = nlev, ess = 1))
      means <- bida:::posterior_mean(bida)
      tmp <- vapply(seq_along(means)[!dindx], 
                    function(x) mean((means[[x]]-pdos[[x]])**2), numeric(1))
      mse[[r, NN, m]] <- tmp
      taus[[r, NN, m]] <- vapply(means[!dindx], bida:::avg_jsd_array, numeric(1))
    }
  }
}
```

### Results
#### MSE of posterior mean of intervention distribution params
```{r}
df <- cbind(expand.grid(dims)[!sapply(mse, is.null), ], do.call("rbind", mse))
df <- cbind(expand.grid(dims)[!sapply(mse, is.null), ], sqrt(vapply(mse, mean, numeric(1))))

df %>% 
  tidyr::pivot_longer(-c(1:3)) %>% 
  mutate(N = factor(N, sort(unique(N)))) %>% 
  ggplot(aes(N, value, color = model)) +
  #facet_grid(, scales = "free") +
  theme_minimal() +
  theme(legend.position = "top",
        legend.title=element_blank()) +
  geom_boxplot() +
  ylab("Root-squared mean errors (RMSE)") +
  xlab("Sample size") 
```

#### Precision recall
```{r message=FALSE, warning=FALSE}
compute_prec_recall <- function(x, y) {
  indx <- order(x, decreasing = TRUE)
  tp  <- cumsum(y[indx])
  cbind(indx = indx, TPR = tp/sum(y), PPV = tp/seq_along(y))
}

# edge probs
par <- expand.grid(dimnames(edgep[,,-3]))
tmp <- lapply(edgep[,,-3], function(x) compute_prec_recall(unlist(x), dag))
indx <- rep(seq_len(nrow(par)), each = n**2)
df  <- cbind(par[indx, ], do.call("rbind", tmp))

df %>% 
  ggplot(aes(TPR, PPV, color = r)) +
  facet_grid(N~model) +
  geom_hline(aes(yintercept = mean(dag)), size = .2) +
  geom_line() +
  ylab("Precision") +
  xlab("True positive rate") +
  ggtitle("Edge probabilities") 

# ancestor relation probs
par <- expand.grid(dimnames(ancp[,,-3]))
tmp <- lapply(ancp[,,-3], function(x) compute_prec_recall(unlist(x), dmat))
indx <- rep(seq_len(nrow(par)), each = n**2)
df  <- cbind(par[indx, ], do.call("rbind", tmp))

df %>% 
  ggplot(aes(TPR, PPV, color = r)) +
  facet_grid(N~model) +
  geom_hline(aes(yintercept = mean(dmat)), size = .2) +
  geom_line() +
  theme_minimal() +
  theme(legend.position = "none") +
  ylab("Precision") +
  xlab("True positive rate") +
  ggtitle("Ancestor relation probabilities") 

# # Strong effects
par <- expand.grid(dimnames(taus))
tmp <- lapply(taus, function(x) compute_prec_recall(unlist(x), strong_effects))
indx <- rep(seq_len(nrow(par)), each = n*(n-1))
df  <- cbind(par[indx, ], do.call("rbind", tmp))

# df %>%
#   ggplot(aes(TPR, PPV, color = r)) +
#   facet_grid(N~model) +
#   geom_hline(aes(yintercept = mean(strong_effects)), size = .2) +
#   geom_line() +
#   theme_minimal() +
#   theme(legend.position = "none") +
#   ylab("Precision") +
#   xlab("True positive rate") +
#   ggtitle("Top 20th percentile strongest effect")

```

```{r}

# avg ppv 
avgppv <- function(x, which_true) {
  pp  <- (length(x)-rank(x) +1)[which_true]
  ppv <- seq_along(pp)/sort(pp)
  mean(ppv)
}

tmp <- lapply(taus, function(x) avgppv(x, which(strong_effects == 1)))
df  <- cbind(expand.grid(dimnames(taus)), value = do.call("rbind", tmp)) 


df %>% 
  group_by(N, model) %>% 
  summarize(mean = mean(value),
            sd   = sd(value)) %>% 
  tidyr::pivot_longer(c("mean", "sd")) %>% 
  mutate(value = as.character(round(value, 2)), 
         value = ifelse(name == "sd", paste0("(", value, ")"), value)) %>% 
  tidyr::pivot_wider(names_from = "model") -> tab

tab %>% 
  select(-name) %>% 
  knitr::kable(digits = c(2))


# positive effects
tmp <- lapply(taus, function(x) avgppv(x, which(dmat[!dindx] == 1)))
df  <- cbind(expand.grid(dimnames(taus)), value = do.call("rbind", tmp)) 


df %>% 
  group_by(N, model) %>% 
  summarize(mean = mean(value),
            sd   = sd(value)) %>% 
  tidyr::pivot_longer(c("mean", "sd")) %>% 
  mutate(value = as.character(round(value, 2)), 
         value = ifelse(name == "sd", paste0("(", value, ")"), value)) %>% 
  tidyr::pivot_wider(names_from = "model") -> tab

tab %>% 
  select(-name) %>% 
  knitr::kable(digits = c(2))

```

