---
title: 'Toy example: sample LDAGs'
author: "Vera Kvisgaard"
date: "2024-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(Rgraphviz)
library(pcalg)
library(BiDAG)
here::i_am("./rmd/sample_ldags_with_BiDAG.Rmd")
invisible(sapply(list.files(here::here("./R/"), ".R", full.names = T), source))
```


```{r}

dag <- rbind(L  = c(0, 1, 1, 0, 1),
             Z1 = c(0, 0, 0, 1, 0),
             Z2 = c(0, 0, 0, 1, 0),
             X  = c(0, 0, 0, 0, 1),
             Y  = rep(0, 5))
colnames(dag) <- rownames(dag)
n <- ncol(dag)
nlev <- rep(2, n)

labels <- matrix(list(), n, n)
labels[[2, 4]] <- rbind(0)  # Z1 -> X
labels[[3, 4]] <- rbind(0)  # Z2 -> X
labels[[4, 5]] <- rbind(0)  # X -> Y

# plot
g <- as(dag, "graphNEL")
indx <- sapply(labels[dag == 1], is.null)
prettify_lab <- function(x) {
  tmp <- apply(x, 1, function(y) sprintf("(%s)",  paste(y, collapse = ",")))
  sprintf("{%s}", paste(gsub("NA", "*", tmp), collapse = ","))
}

lab <- sapply(labels[dag == 1][!indx], prettify_lab)
fonts <- rep(10, length(lab))
names(lab) <- names(fonts) <- Rgraphviz::edgeNames(g)[!indx]
Rgraphviz::plot(g, edgeAttrs = list(label = lab, fontsize = fonts))
```


Draw a distribution over the labeled DAG:
```{r}
set.seed(007)
# store labels as nested list
labels <- lapply(seq_len(n), function(i) labels[which(dag[, i] == 1), i])
bn <- rand_bn(dag, labels, nlev, alpha = 1)
bn
```


Define a score-function that can be used with `BiDAG`. 
Need to assign this function to the namespace of the package, otherwise an internal function is called.
```{r}

# wrapper function
usrDAGcorescore <- function(j, parentnodes, n, scorepar) {
  score_from_lookup(scorepar$data, scorepar$Cvec, j, parentnodes, scorepar$tab)
}

# assign to namespace
assignInNamespace("usrDAGcorescore", usrDAGcorescore, ns = "BiDAG")

```

## Check convergence of MCMC-chain. 

```{r}
data <- bida:::sample_data_from_bn(bn, 100)
```


Specify a `BiDAG::scoreparameters` for LDAGs and standard (discrete) DAGs.
```{r}
get_scorepar <- function(m, data, nlev, tab = NULL) {
 if (m == "DAG") {
      scorepar <- scoreparameters("bdecat", data = as.data.frame(data), bdecatpar = list(chi = 1, edgepf = 1))
    } else {
      if (is.null(tab)) {
        tab <- init_lookup_scoretable(ncol(data), list(ess = 1, kappa = .5))
      }
      scorepar <- scoreparameters("usr", data = data, usrpar = list(pctesttype = "bdecat"))
      scorepar$Cvec <- nlev
      scorepar$tab <- tab
    }
  return(scorepar)
}
```

```{r}

arp_from_MCMCchain <- function(MCMCchain) {
  ln <- length(MCMCchain$traceadd$incidence)
  dmats <- lapply(MCMCchain$traceadd$incidence[seq.int(ceiling(.2*ln), ln)], function(dag) bida:::descendants(as.matrix(dag)))
  Reduce("+", dmats)/length(dmats)
}
```


Run MCMC using the three algorithms `"order", "orderIter", "partition"` twice, and compare the resulting edge probabilities. True edges are marked in blue.
```{r}

m <- "DAG"
indx <- bnlearn::amat(bn) == 1
indx <- bida:::descendants(bn) == 1
par(mfrow = c(2, 2))
tab <- init_lookup_scoretable(ncol(data), list(ess = 1, kappa = .5))

for (m in c("DAG", "LDAG")) {
  edgeps <- arps <- list(order = list(), part = list())
  for (i in c(1, 2)) {
    cat(sprintf("\nRun partitionMCMC to sample %ss, iteration %s\n", m, i))
    tic <- Sys.time()
    scorepar <- get_scorepar(m, data, nlev, tab)
    iterfit  <- learnBN(scorepar, "orderIter", verbose = TRUE, scoreout = TRUE, hardlimit = 5)
    ordfit   <- sampleBN(scorepar, "order", scoretable = getSpace(iterfit))
    partfit  <- sampleBN(scorepar, "partition", scoretable = getSpace(iterfit))
    edgeps$order[[i]]  <- arp_from_MCMCchain(ordfit) #edgep(ordfit) 
    edgeps$part[[i]]   <- arp_from_MCMCchain(partfit) #edgep(partfit) 
    print(Sys.time()-tic)
  }
  
  for(algo in names(edgeps)) {
      plot(edgeps[[algo]][[1]], edgeps[[algo]][[2]], main = paste0(m, "+", algo))
      points(edgeps[[algo]][[1]][indx], edgeps[[algo]][[2]][indx], pch = 19, col = "blue")
      abline(a = 0, b = 1, col = "red")
      
      # plot(0, 0, xlim = c(0, 1), ylim = c(0, 1), xlab = "Recall", ylab = "Precision")
      # for (i in 1:length(edgeps[[algo]])) {
      #   o <- order(edgeps[[algo]][[i]], decreasing = T)
      #   tp <- cumsum(indx[o])
      #   lines(tp/sum(indx), tp/seq_along(tp))
      # }
  } 
}


```

