---
title: "test_optimize_partition"
author: "Vera Kvisgaard"
date: "2024-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold")
```

```{r, results = "hide"}
library(pcalg) # for as(adjmat, "graphNEL")
library(ldags)
here::i_am("./rmd/test_optimize_partition.Rmd")
```

## Test optmization routines
Write a routine that, for a given frequency table, compares the partitions and local scores resulting from each optimization routine `methods = c("tree", "ldag", "part")`.
```{r}
tester <- function(counts, levels, methods = c("tree", "ldag", "part"), regular = F) {
  df <- cbind(expand.grid(levels), n = counts)
  scores <- setNames(numeric(length(methods)), methods)
  for (method in methods) {
    res <- optimize_partition(counts, levels, ess = 1, method, regular = regular)
    df[[method]] <- get_parts(res$partition)
    scores[method] <- sum(res$scores)
  }
  df[["none"]] <- seq_len(nrow(counts))
  scores["none"] <- sum(compute_local_bdeu_score(counts, 1))
  return(list(df, scores))
}
```

### Binary split
Compare partitions for a frequency table with a binary split of the outcome space, where the outcome variable is independent of one variable:
```{r}
levels <- rep(list(0:1), 2)
counts <- cbind(10, c(10, 10, 100, 100))
tester(counts, levels)
```
### V-structure
Compare partitions given a frequency table corresponding to a v-structure with labels on both parents.
```{r}
levels <- rep(list(0:1), 2)
counts <- cbind(10, c(10, 100, 100, 100))
tester(counts, levels)
```
The `tree`-partition returns only one region, i.e. the no-parent CPT, as the score is not improved by splitting on any of the parent variables for the given frequency table. 
```{r}
cat("Score, P(Y|X1, X2):")
compute_local_bdeu_score(counts, ess = 1, r = 2, q = 4, s = 1)
cat("Score, P(Y|X2):")
compute_local_bdeu_score(rowsum(counts, c(1, 1, 2, 2)), ess = 1, r = 2, q = 4, s = c(2, 2))
cat("Score P(Y):")
compute_local_bdeu_score(matrix(colSums(counts), 1), ess = 1, r = 2, q = 4, s = 4)
```
The `make_regular` option of the `optimize_partition` function forces a partition that do not implicitly encode any conditional independencies.
```{r}
tester(counts, levels, regular = T)
```



In the following example, also the `tree` procedure captures the correct local structure.
```{r}
levels <- rep(list(0:1), 2)
counts <- cbind(c(100, 10, 10, 10), c(10, 100, 100, 100))
tester(counts, levels)
```

```{r}
cat("Score, P(Y|X1, X2):")
sum(famscore_bdeu_byrow(counts, ess = 1, r = 2, q = 4, s = 1))
cat("Score, P(Y|X2):")
sum(famscore_bdeu_byrow(rowsum(counts, c(1, 1, 2, 2)), ess = 1, r = 2, q = 4, s = c(2, 2)))
cat("Score P(Y):")
famscore_bdeu_byrow(matrix(colSums(counts), 1), ess = 1, r = 2, q = 4, s = 4)
```


# Compare optimized partitions and search-through-labels routine coincide
```{r}

levels <- rep(list(0:1), 3)
nlev <- lengths(levels)
ess <- 1

for (i in 1:100) {
  #cat("\niter:", i, "\n")
  counts <- matrix(rgamma(2**(3+1), shape = 1), ncol = 2)
  fits <- list()

  fits$part <- optimize_partition(counts, levels, ess, "part", regular = F, F)
  fits$ldag <- optimize_partition(counts, levels, ess, "ldag", regular = F, F)

  partitions <- sapply(lapply(fits, "[[", "partition"), get_parts)
  
  
  # stop and print output if part and ldag do not coincide
  if (!all(partitions[, "part"] == partitions[, "ldag"])) {
    cbind(expand.grid(levels), partitions)
    
    cat("\npart:")
    fits$part <- optimize_partition(counts, levels, ess, "part", regular = F, T)
    cat("\nldag:")
    fits$ldag <- optimize_partition(counts, levels, ess, "ldag", regular = F, T)
    
    is_CSI_consistent(fits$part$partition, nlev)
    is_regular(fits$part$partition, nlev)
    
    is_CSI_consistent(fits$ldag$partition, nlev)
    is_regular(fits$ldag$partition, nlev)
    break
  }
}


```



```{r}
bn   <- readRDS(here::here("./data/LDAG10.rds"))
dag  <- bnlearn::amat(bn)
nlev <- rep(2, 10)  
levels <- lapply(nlev-1, seq.int, from = 0)

data <- bida:::sample_data_from_bn(bn, 10**6)

for (i in seq_along(bn)) {
  par <- which(dag[, i] == 1)
  if (length(par) < 2) next 
  
  parts  <- list()
  cpt <- aperm(bn[[i]]$prob, c(seq_along(par)+1, 1))
  dim(cpt) <- c(prod(nlev[par]), nlev[i]) 
  parts$true <- match(cpt[, 1], cpt[, 1]) 
  
  counts <- compute_freq_table(data, nlev, i, par)
  fit    <- optimize_partition(counts, levels[par], 1, "tree", verbose = F)
  parts$tree  <- get_parts(fit$partition)
  
  fit    <- optimize_partition(counts, levels[par], 1, "ldag", verbose = F)
  parts$ldag  <- get_parts(fit$partition)
  
  cat("\nVariable:", i, "True partition:", parts$true, "\n")
  tabs <- lapply(parts[-1], function(x) table(parts$true, x))
  print(lapply(tabs, function(x) colSums(x > 0)))
  
  #print(data.frame(cpt[, 1], parts))
}            
```


```{r}

i <- 3
par <- which(dag[, i] == 1)
counts <- compute_freq_table(data, nlev, i, par)
microbenchmark::microbenchmark(optimize_partition(counts, levels[par], 1, "part"),
                               optimize_partition(counts, levels[par], 1, "ldag"))

i <- 10
par <- which(dag[, i] == 1)
counts <- compute_freq_table(data, nlev, i, par)
microbenchmark::microbenchmark(optimize_partition(counts, levels[par], 1, "part"),
                               optimize_partition(counts, levels[par], 1, "ldag"))
```




