---
title: "test_optimize_partition"
author: "Vera Kvisgaard"
date: "2024-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      results = "hold")
```

```{r, results = "hide"}
library(pcalg) # for as(adjmat, "graphNEL")
here::i_am("./rmd/test_optimize_partition.Rmd")
sapply(list.files(here::here("./R/"), ".R", full.names = T), source, echo = F)
```

## Test optmization routines
Write a routine that, for a given frequency table, compares the partitions and local scores resulting from each optimization routine `methods = c("tree", "ldag", "part")`.
```{r}
tester <- function(counts, levels, methods = c("tree", "ldag", "part")) {
  df <- cbind(expand.grid(levels), n = counts)
  scores <- setNames(numeric(length(methods)), methods)
  for (method in methods) {
    res <- ldags:::optimize_partition(counts, levels, ess = 1, method, make_regular = F)
    df[[method]] <- ldags:::unlist_partition(res$partition)
    scores[method] <- sum(res$scores)
  }
  return(list(df, scores))
}
```

### Binary split
Compare partitions for a frequency table with a binary split of the outcome space, where the outcome variable is independent of one variable:
```{r}
levels <- rep(list(0:1), 2)
counts <- cbind(10, c(10, 10, 100, 100))
tester(counts, levels)
```
### V-structure
Compare partitions given a frequency table corresponding to a v-structure with labels on both parents.
```{r}
levels <- rep(list(0:1), 2)
counts <- cbind(10, c(10, 100, 100, 100))
tester(counts, levels)
```
The `tree`-partition returns only one region, i.e. the no-parent CPT, as the score is not improved by splitting on any of the parent variables for the given frequency table. 
```{r}
cat("Score, P(Y|X1, X2):")
sum(famscore_bdeu_byrow(counts, ess = 1, r = 2, q = 4, s = 1))
cat("Score, P(Y|X2):")
sum(famscore_bdeu_byrow(rowsum(counts, c(1, 1, 2, 2)), ess = 1, r = 2, q = 4, s = c(2, 2)))
cat("Score P(Y):")
famscore_bdeu_byrow(matrix(colSums(counts), 1), ess = 1, r = 2, q = 4, s = 4)
```

In the following example, also the `tree` procedure captures the correct local structure.
```{r}
levels <- rep(list(0:1), 2)
counts <- cbind(c(100, 10, 10, 10), c(10, 100, 100, 100))
tester(counts, levels)
```
```{r}
cat("Score, P(Y|X1, X2):")
sum(famscore_bdeu_byrow(counts, ess = 1, r = 2, q = 4, s = 1))
cat("Score, P(Y|X2):")
sum(famscore_bdeu_byrow(rowsum(counts, c(1, 1, 2, 2)), ess = 1, r = 2, q = 4, s = c(2, 2)))
cat("Score P(Y):")
famscore_bdeu_byrow(matrix(colSums(counts), 1), ess = 1, r = 2, q = 4, s = 4)
```

# Test score table 
Test the `score_from_lookup` function. For a given node `j` and parent set `parentnodes` this function search in a `rlang::env` list-object `lookup` for the score. If the score is not already computed, it is computed and added to `lookup`. 
```{r}

n <- 5
levels <- rep(list(0:1), n)
data <- sapply(levels, sample, size = 100, replace = T)
tab <- ldags:::init_lookup_scoretable(n, list(ess = 1))

cat("\nEmpty parent set (no CPT-partition-optimization):")
j <- 1
parentnodes <- vector("integer")
ldags:::score_from_lookup(data, levels, lengths(levels), j, parentnodes, method = "tree", lookup = tab)
ls.str(tab)
```

```{r}

cat("\nOne parent (no CPT-partition-optimization):")
j <- 1
parentnodes <- 2
ldags:::score_from_lookup(data, levels, lengths(levels), j, parentnodes, method = "tree", lookup = tab)
ls.str(tab)
```

```{r}
cat("\nMultiple parents:")
j <- 1
parentnodes <- 2:5
ldags:::score_from_lookup(data, levels, lengths(levels), j, parentnodes, method = "tree", lookup = tab)
ls.str(tab)
```

# Score equivalence 
```{r}
n <- 3
levels <- rep(list(0:1), n)
data <- replicate(n, sample(0:1, 100, T))

```

```{r}

dag <- matrix(0, n, n)
dag[upper.tri(dag)] <- 1
colnames(dag) <- rownames(dag) <- c("Z", "X", "Y")
labels <- list(NULL, NULL, list(NULL, 0))
cpts <- lapply(1:n,function(i) cpt_from_dag(dag, levels, labels, i))
for (i in 1:n) {
  cpts[[i]]$counts <- compute_freq_table_from_cpt(cpts[[i]], data)
  cpts[[i]]$score  <- compute_local_score_from_cpt(cpts[[i]], ess = 1)
}
sum(sapply(cpts, "[[", "score"))
  
dag["X", "Y"] <- 0
dag["Y", "X"] <- 1
labels <- list(NULL, list(NULL, 0), NULL)
cpts <- lapply(1:n,function(i) cpt_from_dag(dag, levels, labels, i))
for (i in 1:n) {
  cpts[[i]]$counts <- compute_freq_table_from_cpt(cpts[[i]], data)
  cpts[[i]]$score  <- compute_local_score_from_cpt(cpts[[i]], ess = 1)
}
sum(sapply(cpts, "[[", "score"))

dag["Y", "X"] <- 0
labels <- NULL
cpts <- lapply(1:n,function(i) cpt_from_dag(dag, levels, labels, i))
for (i in 1:n) {
  cpts[[i]]$counts <- compute_freq_table_from_cpt(cpts[[i]], data)
  cpts[[i]]$score  <- compute_local_score_from_cpt(cpts[[i]], ess = 1)
}
sum(sapply(cpts, "[[", "score"))
```


# Test BiDAG
```{r}
dag <- rbind(L  = c(0, 1, 1, 0, 1),
             Z1 = c(0, 0, 0, 1, 0),
             Z2 = c(0, 0, 0, 1, 0),
             X  = c(0, 0, 0, 0, 1),
             Y  = rep(0, 5))
colnames(dag) <- rownames(dag)
n <- ncol(dag)
levels <- rep(list(0:1), n)
nlev <- lengths(levels)

labmat <- matrix(list(), n, n)
labmat[[2, 4]] <- rbind(0)  # Z1 -> X
labmat[[3, 4]] <- rbind(0)  # Z2 -> X
labmat[[4, 5]] <- rbind(0)  # X -> Y

# plot
g <- as(dag, "graphNEL")
indx <- sapply(labmat[dag == 1], is.null)
prettify_lab <- function(x, from, to) {

}

lab <- c()
for (i in seq_len(ncol(dag))) {
  for (j in seq_len(ncol(dag))) {
    if (is.null(labmat[[i, j]])) next
    tmp <- apply(labmat[[i, j]], 1, function(y) sprintf("(%s)",  paste(y, collapse = ",")))
    tmp <- sprintf("{%s}", paste(gsub("NA", "*", tmp), collapse = ","))
    names(tmp) <- sprintf("%s~%s", colnames(dag)[i], colnames(dag)[j])
    lab  <- append(lab, tmp)
  }
}
fonts <- rep(10, length(lab))
names(fonts) <- names(lab)
Rgraphviz::plot(g, edgeAttrs = list(label = lab, fontsize = fonts))
```


Draw a distribution over the labeled DAG:
```{r}
set.seed(007)
partitions <- vector("list", ncol(dag))
for (i in 1:ncol(dag)) {
  pa   <- which(dag[, i] == 1)
  if (length(pa) > 1) {
     labs <- labmat[pa, i]
     if (any(lengths(labs) > 0)) {
       partitions[[i]] <- labels_to_partition(labs, levels[pa], enumerate = T)
     }
  }
}
nlev <- rep(2, ncol(dag))
bn <- rand_bn(dag, partitions, nlev, alpha = 1)
bn
```


```{r}
library(BiDAG)
data <- bida:::sample_data_from_bn(bn, 1000)
scorepar <- scoreparameters("usr", 
                             data = data,
                             usrpar = list(pctesttype = "bdecat"))

scorepar$Cvec <- nlev
scorepar$levels <- lapply(scorepar$Cvec-1, seq.int, from = 0)
scorepar$lookup  <- init_lookup_scoretable(n, list(ess = 1))
ls.str(scorepar$lookup)

# wrapper function
usrDAGcorescore <- function(j, parentnodes, n, scorepar) {
  score_from_lookup(scorepar$data, 
                    levels = scorepar$levels, 
                    nlev = scorepar$Cvec, 
                    j, 
                    parentnodes, 
                    method = scorepar$optPartMethod, 
                    lookup = scorepar$lookup)
}

# assign to namespace
assignInNamespace("usrDAGcorescore", usrDAGcorescore, ns = "BiDAG")

```

```{r}
for (optPartMethod in c("tree", "ldag", "part")){
  scorepar$optPartMethod <- optPartMethod
  scorepar$tab[[optPartMethod]] <- rep(list(list()), ncol(dag))
  print(usrDAGcorescore(1, integer(0), ncol(dag), scorepar))
  print(usrDAGcorescore(1, 2, ncol(dag), scorepar))
  print(usrDAGcorescore(1, 2:3, ncol(dag), scorepar))
}

```

```{r}
res <- list()
for (optPartMethod in c("ldag", "tree")) {
  cat("\n", optPartMethod, "\n")
  scorepar$optPartMethod <- optPartMethod
  scorepar$lookup[[optPartMethod]] <- rep(list(list()), ncol(dag))
  res[[optPartMethod]] <- sampleBN(scorepar, algo = "orderIter", verbose = T, hardlimit = 3)
  
  #tryCatch(sampleBN(scorepar, algo = "orderIter", verbose = T, hardlimit = 3), 
  #         warning = function(cond) stop())
}
```





